Report for Qualifying Exam

### 2024-08-23 src/sections/conclusion.tex
changes in sec 4.5, conclusion, front page

\section{Conclusion}\label{sec:}
In this report, we explored the mechanisms by which words change their meanings over time and how this process can be leveraged to identify data voids.
We examined two key studies to gain insights into this phenomenon.

The first study by Kulkarni et al.\ (2015) presents a robust framework for detecting significant shifts in word usage over time.
Their methodology involves constructing time series for individual words based on frequency, syntactic, and distributional methods, followed by the application of a change point detection algorithm using the Mean Shift model.
This approach effectively identifies moments in time where notable changes in word usage or meaning occur, as illustrated by examples like the evolving meanings of “tape” and “apple,” and the contextual shift of “sandy” post-Hurricane Sandy.

The second study by Hamilton et al.\ (2016) advances the study of semantic change through the development and application of various word embedding models — PPMI, SVD, and SGNS — to track the evolution of word meanings over time.
By aligning embeddings across different periods and employing metrics like Spearman correlation and Procrustes alignment,
they uncover two important statistical laws: the Law of Conformity, which indicates that more frequent words undergo smaller semantic changes,
and the Law of Innovation, which shows that words with higher polysemy are more likely to experience significant shifts in meaning.
Their findings provide valuable tools and insights into the dynamics of language evolution.

Through these discussions, we have gained a deeper understanding of how semantic change can be detected and analyzed,
offering powerful tools for identifying data voids in evolving linguistic landscapes.

>>>>

\section{Conclusion}\label{sec:}
In this report, we explored the mechanisms by which words change their meanings over time and how this process can be leveraged to identify \emph{data voids}.
We examined two key studies to gain insights into this phenomenon.

Kulkarni et al.\ (2015) introduced the concept of creating \emph{word embeddings} to measure the context of a word over time.
By constructing these embeddings and applying \emph{change point detection}, they provided a powerful method for identifying significant shifts in word meanings.
This approach can be instrumental in detecting data voids, as abrupt changes in word usage may signal emerging topics or areas where existing data is insufficient or outdated.

Hamilton et al.\ (2016) built on this foundation by applying variations of the word2vec model, particularly using the Skip-Gram with Negative Sampling (SGNS) method,
to track the evolution of word meanings across different time periods.
Hamilton’s unique contribution lies in the formulation of statistical laws governing semantic change, such as the Law of Conformity and the Law of Innovation.
The \emph{Law of Conformity} indicates that more frequent words tend to undergo smaller semantic changes,
suggesting that less frequent words may be more prone to shifts that could lead to data voids.
Meanwhile, the \emph{Law of Innovation} reveals that words with higher polysemy are more likely to experience significant shifts in meaning.
This insight is crucial for identifying data voids, as words with multiple meanings that rapidly evolve may create gaps in data coverage,
highlighting areas where new data collection or analysis is needed to keep pace with linguistic changes.

Monitoring such shifts can help identify areas where existing data may no longer capture the full spectrum of meanings.

### 2024-08-23 src/sections/paper_kulkarni.tex
changes in sec 4.5, conclusion, front page

\para{Change point algorithm.}
After constructing time series data for each word in the corpus—whether using frequency-based, syntactic, or distributional methods (\Cref{subsec:constructing-word-embeddings}), the first step is to normalize the time series data for the word $w$.
After normalization, the algorithm computes a mean shift series, denoted as $K(Z(w))$.
The \emph{mean shift} is defined as the difference in the mean of a time series at different time points.
The algorithm calculates this shift by segmenting the time series around a pivot point, denoted as time point $j$, and comparing the means of the segments before and after this point.
To assess the significance of the computed mean shifts, the algorithm employs a \emph{bootstrapping} method.
The algorithm uses bootstrap sampling to ensure that the detected change point is not just due to random noise.
It generates multiple synthetic time series by randomly resampling the original time series, denoted as $BS$.
These resampled sets serve as a reference to evaluate whether the detected shifts in the real data are significant or not.
This is done by repeatedly drawing samples from the normalized data $Z(w)$ until the number of samples equals $B$, which is predefined.
For each sample, the algorithm calculates a p-value for each time point $i$.
This p-value indicates how likely it is that the observed change at that time point is significant.
The algorithm identifies potential change points by looking for times $j$ where the Z-Score $Z_j(w)$ is greater than or equal to the threshold $\gamma$.
It then checks the p-values for these change points and selects the one with the smallest p-value, indicating the most significant change.
The algorithm returns the p-value and the estimated change point (ECP).
This information helps understand exactly when the meaning or usage of the word has shifted over time.
%\Cref{fig:kulkarni-changepoint} illustrates the important aspects of the algorithm.

>>>>

\para{Change point algorithm.}
After constructing time series data for each word in the corpus—whether using frequency-based, syntactic, or distributional methods (\Cref{subsec:constructing-word-embeddings}), the first step is to normalize the time series data for the word $w$.
After normalization, the algorithm computes a mean shift series, denoted as $K(Z(w))$.
The \emph{mean shift} is defined as the difference in the mean of a time series at different time points.
The algorithm calculates this shift by segmenting the time series around a pivot point, denoted as time point $j$, and comparing the means of the segments before and after this point.
To assess the significance of the computed mean shifts, the algorithm employs a \emph{bootstrapping} method.
The algorithm uses bootstrap samplingto estimate how likely an observed shift is, under the assumption that no real change has occurred (the null hypothesis).
It generates multiple time series by randomly resampling the original time series, denoted as $BS$.
This is done by repeatedly drawing samples from the normalized data $Z(w)$ until the number of samples equals $B$, which is predefined.
The p-value is then calculated by comparing the observed mean shift to the distribution of mean shifts from the bootstrapped samples.
It reflects the proportion of times that the mean shift from the random (permuted) samples exceeds the observed mean shift.
A low p-value indicates that the observed shift is unlikely to be due to random fluctuations alone.
The algorithm identifies potential change points by looking for times $j$ where the Z-Score $Z_j(w)$ is greater than or equal to the threshold $\gamma$,
which is set to the 96th percentile, to filter out less significant potential change points.
It then checks the p-values for these change points and selects the one with the smallest p-value, indicating the most significant change.
The algorithm returns the p-value and the estimated change point (ECP).
This information helps understand exactly when the meaning or usage of the word has shifted over time.
%\Cref{fig:kulkarni-changepoint} illustrates the important aspects of the algorithm.
